Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex
Forgot to change sim to preceq
rest of algebra

\begin{definition} The \textsf{Mult}
and \textsf{Approx} are parametric labelled variant (sum) types 
with injection function given by their definition:
%
\begin{align*}
\textsf{Mult} \, a & = \textsf{mult} \, a \,\mid\, \textsf{only} \, a \\
\textsf{Approx} \, a & = \textsf{exact} \, a \,\mid\, \textsf{lower} \, a \,\mid\,
\textsf{upper} \, a \,\mid\, \textsf{both} \, a \, a
\end{align*}
\ie{}, $\mathsf{lower}$ is an injection $\mathsf{lower} : a \rightarrow \mathsf{Approx} \, a$, and so
on. These are used for labelled models, and other components later in
Section~\ref{sec:analysis} with additional information.
\label{def:mult-and-approx}
\end{definition}
%

\begin{figure}
\input{spatial-model.tex}
\caption{\textit{spec} model, 
$\interp{-}_n : \textit{spec} \rightarrow \textsf{Mult}(\textsf{Approx}(\mathcal{P}(\domainVal{}^n)))$}
\label{fig:spatial-model}
\end{figure}

@@

Given the model defined in this section we can
consider soundness for the equational theory of
specifications and the notion of spatial approximation
on specifications. The proofs are given in our supplement.

\begin{theorem}[Soundness of equational theory]
\[
\overline{\texttt{v} : S}\equiv \overline{\texttt{u} : T}
\; \Rightarrow \;
\interp{\overline{\texttt{v} : S}} = \interp{\overline{\texttt{u} : T}}
\]
\end{theorem}

@@

\item Sort each permutation set into an ordered list, based on the ordering:
\begin{equation*}
  \vect{x} \leq \vect{y} =
      \exists i \, . \; \vect{x}_{i,1} \leq \vect{y}_{i,1} \; \wedge \;
        (i = n \vee \vect{x}_{i+1:n} = \vect{y}_{i+1:n})
\end{equation*}
%
This yields:
%%
\begin{align*}
\{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{0}{1}{0}{1}] )
\\
&(\pi^2_2, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{0}{1}{0}{1}])\}
\end{align*}
%%
\item Fold each list pairwise by the following partial operation
 $\bullet$ which coalesces contiguous regions:
%
\begin{align*}
& (\vect{x}^L,\vect{x}^U) \bullet (\vect{y}^L,\vect{y}^U) \\
= &
\begin{cases}
(\vect{x}^L, \vect{y}^U) & \vect{x}^U_1 + 1 = \vect{y}^L_1 \; \wedge \;
\vect{x}_{2:n} = \vect{y}_{2:n}) \\
\bot  & \textit{otherwise}
\end{cases}
\end{align*}
For our example, this yields:
%
\begin{align*}
Q(v) = \{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1}
]) \\
&(\pi^2_2, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1}])\}
\end{align*}
%
\item Un-permute and union together, \ie{},
%
\[
U(v) = \bigcup \{[\pi \vect{x}^L, \pi \vect{x}^U]
 \mid [\vect{x}^L, \vect{x}^U] \leftarrow S, (\pi, S) \leftarrow Q(v)\}
\]
Which yields:
%
\begin{align*}
U(\texttt{a}) =
\{\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{-1}{0}{1},
\stwo{1}{0}{1}{0}\}
\end{align*}
%
\item Finally, filter by the region containment predicate $\containedin$, that
  is, if any region is contained within another then remove the
  smaller. The $\containedin$ predicate is defined:
%
\begin{align*}
& (\vect{x}^L, \vect{x}^U) \containedin (\vect{y}^L, \vect{y}^U) = \\
& \vect{y}^L_1 \leq \vect{x}^L_1 \wedge \vect{x}^U_1 \leq \vect{y}^U_1
  \wedge (\vect{x}^L_{2:n}, \vect{x}^U_{2:n}) \containedin
  (\vect{y}^L_{2:n}, \vect{y}^U_{2:n})
\end{align*}
For our example this then yields the final result of:
\begin{align*}
\spanOp(M)
= \{\stwo{-1}{0}{1}{0} \stwo{0}{-1}{0}{1}\}
\end{align*}
since $\stwo{0}{1}{0}{1} \sqsubseteq \stwo{0}{-1}{0}{1}$
and $\stwo{1}{0}{1}{0} \sqsubseteq \stwo{-1}{0}{1}{0}$.
\end{enumerate}
For our example, applying \spanOp again yields the same
result, hence we have reached a fixed point.

@@

\newcommand{\bplus}{\operatornamewithlimits{\term{+}}}
\newcommand{\tySum}[1]{#1^{\term{+}}}
\newcommand{\tyProd}[1]{#1^{\term{*}}}
\newcommand{\specDNF}{\textbf{spec}}
\newcommand{\result}[1]{\lfloor{#1}\rceil}

Next, the set of spans $\spanOp(M)$ covering the indexing
space is converted into abstract syntax trees 
for the specification language. This conversion is defined in terms
of an algebra on \textit{spec} syntax in disjunctive-normal form
(which we denote \specDNF{}).

\begin{align*}
\result{\specDNF{}} = \specDNF + ((\specDNF{} + 1) \times
  (\specDNF{} + 1))
\end{align*}

\begin{align*}
\result{\specDNF{}} & ::= \textsf{exact} \, \specDNF{}
\mid \textsf{bounds} ([\specDNF{}], [\specDNF{}]) \\
\end{align*}

The algebra mirrors the shape
of the region combinators for addition and multiplication, 
with operators $\oplus$ and $\odot$ and a constructor for \specDNF{}:
%%
\begin{align*}
\oplus & : \specDNF \rightarrow \specDNF \rightarrow
  \specDNF \\
\odot & : \specDNF \rightarrow \specDNF \rightarrow
        \specDNF \\
\mathsf{toSpec} & : (\mathbb{Z} \cup \{\infty\}) \rightarrow (\mathbb{Z} \cup
  \{\infty\})  \rightarrow \specDNF
\end{align*}
%%
The $\mathsf{toSpec}$ operation takes a pair of lower and upper bounds
values (one-dimensional) and constructs a specification 
which contains a one-dimensional region. 
An entire span is converted into an $n$-dimensional
specification by applying $\mathsf{toSpec}$ to each
pair of lower and upper bounds in a span, per dimension, and multiplying these
specifications by $\odot$. The resulting specifications for
each span are then summed by by $\oplus$, that is, this step is given by:
\begin{equation*}
\textsf{spec}(M) = 
\sum\limits_{\vect{x} \in \spanOp(M)}
\prod\limits_{i \in \{1, \ldots, n\}} \, \mathsf{toSpec}(\vect{x}^L_i, \vect{x}^U_i)
\end{equation*}
(where $\oplus$ is the additive and $\odot$ is the multiplicative used
within $\Sigma$ and $\Pi$ respectively here).
We unpack the definitions of this algebra.

\begin{definition}[Specifications in DNF]
Let $r_1, ..., r_n$ range over elements of the \textit{region}
syntax (see Figure~\ref{fig:syntax}, p.~\pageref{fig:syntax}).
The set \specDNF{} is a subset of syntax trees \textit{spec}
of specifications in disjunctive normal form over regions 
where the disjunct is \term{+} (since its model is in terms of set
union) and the conjunct is \term{*} (since its model is a form of tensor).
That is elements of \specDNF{} are of the form:
%%
\[
(r^1_{1} \; \term{*} \ldots \term{*} \; r^1_n)\; \term{+} \; \ldots 
\term{+} \; (r^m_1 \, \term{*} \; \ldots \term{*} \; r^m_p)
\]
%%
More formally, $\specDNF \subseteq \textit{spec}$, defined:
%%
\begin{align*}
\specDNF{} & ::= [\textit{mult},] \;
\tySum{\specDNF} \\
\tySum{\specDNF} & ::= \tyProd{\specDNF} \; \term{+} \; \tySum{\specDNF} \mid
  \tyProd{\specDNF} \\
\tyProd{\specDNF} & ::= \textit{region} \; \term{*} \; \tyProd{\specDNF} \mid
   \textit{region}
\end{align*}
%%
In the following, $s, t$ range over elements of
\specDNF{}, $\tySum{s}, \tySum{t}$ range over elements of 
$\tySum{\specDNF}$ and $\tyProd{s}, \tyProd{t}$ range over elements
of $\tyProd{\specDNF}$ (for example, $\tyProd{s} = r^1_{1} \; \term{*}
\ldots \term{*} \; r^1_n$).
\end{definition}

\begin{definition} The $\mathsf{toSpec}$ constructor is defined:
%%
%\newcommand{\llower}{\vect{x}^L_i}
%\newcommand{\lupper}{\vect{x}^U_i}
\newcommand{\llower}{l}
\newcommand{\lupper}{u}
\begin{align*}
&\mathsf{toSpec} \; \llower \; \lupper = \\
&\begin{cases}
%% FWD
 \stenFwdS{\lupper}{i}  & \llower = 0 \, \wedge \, \lupper > 0 \\[0.2em]
 \stenFwdSR{\lupper}{i}{\irreflS}  & \llower = 1 \, \wedge \, \lupper > 0 \\[0.2em]
%% BWD
 \stenBwdS{|\llower|}{i}  & \llower < 0 \, \wedge \, \lupper = 0 \\[0.2em]
\stenBwdSR{|\llower|}{i}{\irreflS} & \llower < 0 \, \wedge \, \lupper = -1 \\[0.2em]
%% CEN
 \stenCenS{\lupper}{i}  & \llower < 0 \, \wedge \, \lupper > 0
 \, \wedge \, |\llower| = \lupper \\[0.2em]
 \stenBwdS{|\llower|}{i} & \llower < 0 \, \wedge \, \lupper > 0  \, \wedge \, |\llower| \neq \lupper \\
\quad \texttt{+} \, \stenFwdS{\lupper}{i}  & \qquad\quad \\[0.2em]
% BOUND
 \texttt{atMost}, \stenFwdS{\lupper}{i} & \llower > 1 \\[0.2em]
 \texttt{atMost}, \stenBwdS{|\llower|}{i} \quad & \llower < (-1)
 \\[0.2em]
% ABSOLUTE REP
 \epsilon & \llower = \infty \, \vee \, \lupper = \infty
  \end{cases}
\end{align*}
%%
The first two cases map to \texttt{forward} stencils since the lower
bound is $0$ or $1$, with the depth determined by the upper bound of
the span. If the lower bound is $1$, then the
$0$-offset point is missing so the region is given an
\texttt{irreflexive} attribute. The second two cases are similar for
\texttt{backward} but now the depth of the region is determined by the
absolute of the lower bound. In the fifth-case, where the lower bound
is negative and the upper bound is positive and they have the same
magnitude, then a \texttt{centered} region is constructed with
the depth as the common magnitude. If the magnitudes are not the same
(sixth case) then separated \texttt{forward} and \texttt{backward}
regions are summed, with their corresponding depths.

@@

Finally, in the case where either bound is $\infty$, representing a non
neighbourhood index, then an empty specification is given
 with no regions, indicating that there is not even reflexive
0-offset access in this dimension.
\end{definition}

\begin{lemma}
$\mathsf{toSpec}$ is total (\ie{}, defined for all inputs)
 given the precondition that the lower
bound of the input span is less than or equal to its upper bound.
\end{lemma}

%\newcommand{\mkBig}[2]{\overset{\LARGE{\term{#1}}}{#2}}
\begin{definition} The $\odot$ operation on \specDNF{} is defined
\begin{align*}
& ([\textit{approx}_1] \; [\textit{mult}_1] \; 
\tySum{s})
\odot
([\textit{approx}_2] \; [\textit{mult}_2] \;
\tySum{t}) = \\
& \begin{cases}
\tySum{s} \tySum{\odot} \tySum{t} & \textit{approx}_1 =
\textit{approx_2} = \varepsilon
\\
\
& \textit{approx}_1 = \texttt{atLeast}, \textit{approx}_2 = \varepsilon
\end{cases}
\\
\end{align*}
\begin{align*}
s \tySum{\odot} t & =
 \bplus_{(\tyProd{s_1}, \tyProd{s_2}) \in (s \times
   t)} (\tyProd{s_1} \; \term{*} \; \tyProd{s_2})
%\begin{cases}%
%\end{cases}
\end{align*}
\end{definition}

>>>>

\begin{definition} The \textsf{Mult}
and \textsf{Approx} are parametric labelled variant types 
with injection functions given by their definition:
%
\begin{align*}
\textsf{Mult} \, a & = \textsf{mult} \, a \,\mid\, \textsf{only} \, a \\
\textsf{Approx} \, a & = \textsf{exact} \, a \,\mid\, \textsf{lower} \, a \,\mid\,
\textsf{upper} \, a \,\mid\, \textsf{both} \, a \, a
\end{align*}
\ie{}, $\mathsf{lower}$ is an injection $\mathsf{lower} : a \rightarrow \mathsf{Approx} \, a$, and so
on. The $\textsf{Mult}$ type corresponds to the presence or absence of the
\term{readOnce} modifier as shown in Figure~\ref{fig:spatial-model}. 
The $\textsf{Approx}$ type corresponds to the presence or absence of
the spatial approximation modifier, with \textsf{exact} when there is
no such modifier and \textsf{lower} and \textsf{upper} for
\term{atLeast} and \term{atMost}. Later, it will be useful to model
pairs of lower and upper bounds which is provided by \textsf{both} 
(but which isn't used here).

These are used for labelled models, and other components later in
Section~\ref{sec:analysis} with additional information.
\label{def:mult-and-approx}
\end{definition}
%

\begin{figure}[t]
\input{spatial-model.tex}
\caption{\textit{spec} model, 
$\interp{-}_n : \textit{spec} \rightarrow \textsf{Mult}(\textsf{Approx}(\mathcal{P}(\domainVal{}^n)))$}
\label{fig:spatial-model}
\vspace{-1em}
\end{figure}

@@

Given the model defined in this section we can
consider soundness for the equational theory of
specifications and the notion of spatial approximation. The proofs are given in our supplement.

\begin{theorem}[Soundness of equational theory (Figure~\ref{fig:equations})]
\[
S \equiv T \; \Rightarrow \;
\interp{S}_n = \interp{T}_n
\]
\end{theorem}

@@

\item Sort each permutation set into an ordered list, based on the ordering:
\begin{equation*}
  \vect{x} \leq \vect{y} =
      \exists i \, . \; \vect{x}_{i,1} \leq \vect{y}_{i,1} \; \wedge \;
        (i = n \vee \vect{x}_{i+1:n} = \vect{y}_{i+1:n})
\end{equation*}
%
This yields:
%%
\begin{align*}
\{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{0}{1}{0}{1}] )
\\
&(\pi^2_2, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{0}{1}{0}{1}])\}
\end{align*}
%%
\item Fold each list pairwise by the following partial operation
 $\bullet$ which coalesces contiguous regions:
%
\begin{align*}
\vect{x} \bullet \vect{y}
= &
\begin{cases}
(\vect{x}^L, \vect{y}^U) & \vect{x}^U_1 + 1 = \vect{y}^L_1 \; \wedge \;
\vect{x}_{2:n} = \vect{y}_{2:n}) \\
\bot  & \textit{otherwise}
\end{cases}
\end{align*}
For our example, this yields:
%
\begin{align*}
Q(v) = \{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1}
]) \\
&(\pi^2_2, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1}])\}
\end{align*}
%
\item Un-permute and union together, \ie{},
%
\[
U(v) = \bigcup \{[\pi \vect{x}^L, \pi \vect{x}^U]
 \mid [\vect{x}^L, \vect{x}^U] \leftarrow S, (\pi, S) \leftarrow Q(v)\}
\]
Which yields:
%
\begin{align*}
U(\texttt{a}) =
\{\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{-1}{0}{1},
\stwo{1}{0}{1}{0}\}
\end{align*}
%
\item Finally, filter by the region containment predicate $\containedin$, that
  is, if any region is contained within another then remove the
  smaller. The $\containedin$ predicate is defined:
%
\begin{align*}
& (\vect{x}^L, \vect{x}^U) \containedin (\vect{y}^L, \vect{y}^U) = \\
& \vect{y}^L_1 \leq \vect{x}^L_1 \wedge \vect{x}^U_1 \leq \vect{y}^U_1
  \wedge (\vect{x}^L_{2:n}, \vect{x}^U_{2:n}) \containedin
  (\vect{y}^L_{2:n}, \vect{y}^U_{2:n})
\end{align*}
For our example this then yields the final result of:
\begin{align*}
\spanOp(M)
= \{\stwo{-1}{0}{1}{0} \stwo{0}{-1}{0}{1}\}
\end{align*}
since $\stwo{0}{1}{0}{1} \sqsubseteq \stwo{0}{-1}{0}{1}$
and $\stwo{1}{0}{1}{0} \sqsubseteq \stwo{-1}{0}{1}{0}$.
\end{enumerate}
For our example, applying \spanOp again yields the same
result, hence we have reached a fixed point.

@@

\newcommand{\bplus}{\operatornamewithlimits{\term{\huge{+}}}}
\newcommand{\tySum}[1]{#1^{\term{+}}}
\newcommand{\tyProd}[1]{#1^{\term{*}}}
\newcommand{\specDNF}{\textbf{spec}}

Next, the set of spans $\spanOp(M)$ covering the indexing
space is converted into abstract syntax trees 
of the specification language. This conversion is defined in terms
of an algebra on \textit{spec} syntax in disjunctive-normal form
(which we denote \specDNF{}), and which is wrapped
with the $\textsf{Approx}$ type (see
Definition~\ref{def:mult-and-approx},
p.~\pageref{def:mult-and-approx}), which we write as $\mathsf{A}$
for brevity.  The algebra mirrors the shape
of the region combinators for addition and multiplication, 
with operators $\oplus$ and $\odot$ and a constructor for \specDNF{}:
%%
\begin{align*}
\oplus & : \mathsf{A}(\specDNF) \rightarrow \mathsf{A}(\specDNF) \rightarrow
  \mathsf{A}(\specDNF) \\
\odot & : \mathsf{A}(\specDNF) \rightarrow \mathsf{A}(\specDNF) \rightarrow
        \mathsf{A}(\specDNF) \\
\mathsf{toSpec} & : (\mathbb{Z} \cup \{\infty\}) \rightarrow (\mathbb{Z} \cup
  \{\infty\})  \rightarrow \mathsf{A}(\specDNF)
\end{align*}
%%
The $\mathsf{toSpec}$ operation takes a pair of lower and upper bounds
values (one-dimensional) and constructs a specification 
which contains a one-dimensional region. 
An entire span is converted into an $n$-dimensional
specification by applying $\mathsf{toSpec}$ to each
pair of lower and upper bounds in a span, per dimension, and multiplying these
specifications by $\odot$. The resulting specifications for
each span are then summed by by $\oplus$, that is, this step is given by:
\begin{equation*}
\textsf{spec}(M) = 
\sum\limits_{\vect{x} \in \spanOp(M)}
\prod\limits_{i \in \{1, \ldots, n\}} \, \mathsf{toSpec}(\vect{x}^L_i, \vect{x}^U_i)
\end{equation*}
(where $\oplus$ is the additive and $\odot$ is the multiplicative used
within $\Sigma$ and $\Pi$ respectively here).
We unpack the definitions of this algebra.

\begin{definition}[Specifications in DNF]
Let $r_1, ..., r_n$ range over elements of the \textit{region}
syntax (see Figure~\ref{fig:syntax}, p.~\pageref{fig:syntax}).
The set \specDNF{} is a subset of syntax trees \textit{spec}
of specifications in disjunctive normal form over regions 
where the disjunct is \term{+} (since its model is in terms of set
union) and the conjunct is \term{*} (since its model is a form of tensor).
That is elements of \specDNF{} are of the form:
%%
\[
(r^1_{1} \; \term{*} \ldots \term{*} \; r^1_n)\; \term{+} \; \ldots 
\term{+} \; (r^m_1 \, \term{*} \; \ldots \term{*} \; r^m_p)
\]
%%
More formally, $\specDNF \subseteq (\textit{spec} \cup \{\epsilon\})$, defined:
%%
\begin{align*}
\specDNF{} & ::= \tySum{\specDNF} \mid \epsilon \\
\tySum{\specDNF} & ::= \tyProd{\specDNF} \; \term{+} \; \tySum{\specDNF} \mid
  \tyProd{\specDNF} \\
\tyProd{\specDNF} & ::= \textit{region} \; \term{*} \; \tyProd{\specDNF} \mid
   \textit{region}
\end{align*}
%%
We incldue the empty element $\epsilon$ which is not present in the
syntax. 
In the following, $s, t$ range over elements of
\specDNF{}, $\tySum{s}, \tySum{t}$ range over elements of 
$\tySum{\specDNF}$ and $\tyProd{s}, \tyProd{t}$ range over elements
of $\tyProd{\specDNF}$ (for example, $\tyProd{s} = r^1_{1} \; \term{*}
\ldots \term{*} \; r^1_n$).
\end{definition}

\newcommand{\exactR}{\mathsf{exact}\;}
\newcommand{\exact}[1]{\exactR(#1)}
\newcommand{\upperB}[1]{\mathsf{upper}\;(#1)}

\begin{definition} The $\mathsf{toSpec}$ constructor is defined:
%%
%\newcommand{\llower}{\vect{x}^L_i}
%\newcommand{\lupper}{\vect{x}^U_i}
\newcommand{\llower}{l}
\newcommand{\lupper}{u}
\begin{align*}
&\mathsf{toSpec} \; \llower \; \lupper = \\
&\hspace{-0.4em}\begin{cases}
%% FWD
\exact{\stenFwdS{\lupper}{i}}  & \llower = 0 \, \wedge \, \lupper > 0 \\[0.2em]
\exact{\stenFwdSR{\lupper}{i}{\irreflS}}\!\!\!\! & \llower = 1 \, \wedge \, \lupper > 0 \\[0.2em]
%% BWD
\exact{\stenBwdS{|\llower|}{i}}  & \llower < 0 \, \wedge \, \lupper = 0 \\[0.2em]
\exact{\stenBwdSR{|\llower|}{i}{\irreflS}}\!\!\!\! & \llower < 0 \, \wedge \, \lupper = -1 \\[0.2em]
%% CEN
\exact{\stenCenS{\lupper}{i}}& \llower < 0 \wedge \lupper > 0 \wedge |\llower| = \lupper \\[0.2em]
\exactR(\stenBwdS{|\llower|}{i} & \llower < 0 \wedge \lupper > 0  \wedge |\llower| \neq \lupper \\
\qquad\qquad \texttt{+} \, \stenFwdS{\lupper}{i}) & \\[0.2em]
% BOUND
\upperB{\stenFwdS{\lupper}{i}} & \llower > 1 \\[0.2em]
\upperB{\stenBwdS{|\llower|}{i}} \quad & \llower < (-1)
 \\[0.2em]
% ABSOLUTE REP
\exact{\epsilon} & \llower = \infty \, \vee \, \lupper = \infty
  \end{cases}
\end{align*}
%%
The first two cases map to \texttt{forward} stencils since the lower
bound is $0$ or $1$, with the depth determined by the upper bound of
the span. If the lower bound is $1$, then the
$0$-offset point is missing so the region is given an
\texttt{irreflexive} attribute. The second two cases are similar for
\texttt{backward} but now the depth of the region is determined by the
absolute of the lower bound. In the fifth-case, where the lower bound
is negative and the upper bound is positive and they have the same
magnitude, then a \texttt{centered} region is constructed with
the depth as the common magnitude. If the magnitudes are not the same
(sixth case) then separated \texttt{forward} and \texttt{backward}
regions are summed, with their corresponding depths.

@@

Finally, if either bound is $\infty$, representing a non
neighbourhood index, then an empty specification is given
 with no regions, indicating that there is not even reflexive
0-offset access in this dimension.
\end{definition}

\begin{lemma}
$\mathsf{toSpec}$ is total 
 given the precondition that the lower
bound of the input span is less than or equal to its upper bound.
\end{lemma}

%\newcommand{\mkBig}[2]{\overset{\LARGE{\term{#1}}}{#2}}
\begin{definition} The $\odot$ operation on $\textsf{A}(\specDNF{})$
  is defined in terms of the intermediate $\hat{\odot}$ on \specDNF{}:
%%
\begin{align*}
S \,\hat{\odot}\, \epsilon = S \quad
\epsilon \,\hat{\odot}\, S = S \qquad
S \,\hat{\odot}\, T = \bplus_{(\tyProd{s}, \tyProd{t}) \in (S \times
   T)} (\tyProd{s} \; \term{*} \; \tyProd{t})
\end{align*}
%%
That is, the $\hat{\odot}$ product of two specifications $S$ and $T$ in DNF form 
is the $\term{+}$ sum of all pairwise $\term{*}$ products for every
pair of $\tyProd{spec}$ drawn from $S$ and $T$. The result
is indeed in DNF-form. The operation is
commutative and associative, and sound with respect to our
model (see Lemma~\ref{lem:alg-soundness} below).
The $\hat{\odot}$ operation is then lifted to all combinations of
\textsf{Approx} for $\odot$:
\begin{align*}
\textsf{exact}(S) \odot \textsf{exact}(T) = \; & \textsf{exact}(S
                                              \hat{\odot} T) \\
\textsf{lower}(S) \odot \textsf{lower}(T) = \; & 
      \textsf{lower} \, (S \hat{\odot} T) \\
\textsf{upper}(S) \odot \textsf{upper}(T) = \; &
      \textsf{upper} \, (S \hat{\odot} T) \\
\textsf{lower}(S) \odot \textsf{exact}(T) = \; &
       %\textsf{exact}(T) \odot \textsf{lower}(S) = \;
       \textsf{both} \, (S \hat{\odot} T, \; T) \\
\textsf{upper}(S) \odot \textsf{exact}(T) = \; & 
       %\textsf{exact}(T) \odot \textsf{upper}(S) =
      \; \textsf{both} \, (T, \; S \hat{\odot} T) \\
\textsf{lower}(S) \odot \textsf{both}(T_l,T_u) = \; &
        %\textsf{both}(T_l,T_u) \odot \textsf{lower}(S) = \;
        \textsf{both} \, (S \hat{\odot} T_l, \; T_u) \\
\textsf{upper}(S) \odot \textsf{both}(T_l,T_u) = \; & \textsf{both} \,
                                                      (T_l, \; S \hat{\odot} T_u) \\
\textsf{both}(S_l,S_u) \odot
\textsf{both}(T_l,T_u) & = \textsf{both} (S_l \hat{\odot} T_l, \; S_u
                         \hat{\odot} T_u)
\end{align*}
For brevity we omit the cases where the parameters above are flipped
as $\odot$ is commutative.
\end{definition}

\begin{definition}The $\oplus$ operation on $\textsf{A}(\specDNF)$ 
is defined in a similar way to $\odot$, with $\hat{\oplus}$ on
$\specDNF$ given by:
%%
\begin{equation*}
S \,\hat{\oplus}\, \epsilon = S \quad
\epsilon \,\hat{\oplus}\, S = S \quad
S \, \hat{\oplus} \, T = S \, \term{+} \, T
\end{equation*}
%%
Thus, if neither specification is empty, their $\oplus$ is just
the syntatic sum $\term{+}$. Then $\oplus$ is the lifting of
$\hat{\oplus}$ to boudns with the shape as $\odot$ in the previous
definition.
\end{definition}

\begin{lemma}[$\hat{\odot}$, $\hat{\oplus}$ soundness] $\forall S,
  T \in \specDNF$ where $S \neq \epsilon \wedge T \neq \epsilon$:
\begin{align*}
\interp{S \; \hat{\odot} \; T}_n & = \interp{S}_n \otimes \interp{T}_n
  \\
\interp{S \; \hat{\oplus} \; T}_n & = \interp{S}_n \cup \interp{T}_n
\end{align*}
\label{lem:alg-soundness}
\end{lemma}
%
